A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array

### 2014-10-28 foreground_delay_spectrum_modeling.tex
Minor changes in wording of impact of baseline screening

This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data. Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{We use standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window} as a measure of foreground contamination.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings. We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings. In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied. % Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference. 
The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5). It is notable that short antenna spacings play maximum role in determining foreground contamination. For instance, among antenna spacings with orientations $|\theta_\textrm{b}|\le$~22\fdg 5, discarding those with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%. On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings accounting for $\sim 25$\% loss of data.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
\caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines. Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\vec{b}| \le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded. Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases. Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5). The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|_\textrm{max}\gtrsim 30$~m) has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}

@@

In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum. We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, without significant loss of data.

>>>>

This screening technique can be generalized to optimize between foreground mitigation and loss of sensitivity from discarding data. Figure~\ref{fig:screening} shows how the typical foreground contamination\footnote{Foreground contamination is measured by standard deviation of noiseless $V_{b\tau}(\vec{b},\tau)$ from foregrounds in the {\it EoR window}.} in the {\it EoR window} depends on the orientations and lengths of discarded antenna spacings. We choose antenna spacings oriented eastward to varying degrees of directedness, i.e., -7\fdg 5~$\le\theta_\textrm{b}<$~7\fdg 5 (solid circles), -15\arcdeg~$\le\theta_\textrm{b}<$~15\arcdeg~ (solid squares), and -22\fdg 5~$\le\theta_\textrm{b}<$~22\fdg 5 (solid stars). Among antenna spacings that satisfy these criteria, we discard data from those whose lengths are shorter than $|\vec{b}|_\textrm{max}$ ($x$--axis) and show foreground contamination estimated in the {\it EoR window} from all remaining antenna spacings. In other words, these plots demonstrate the progress in mitigation as orientation and maximum length of discarded antenna spacings are varied. % Foreground contamination estimated from only northward (dashed line) and eastward (dot--dashed line) antenna spacings represent the lower and upper limits, respectively, and are shown for reference. 
The fraction of discarded antenna spacings discarded relative to the total number is shown in dotted lines for different ranges of $\theta_\textrm{b}$. It is seen that foreground contamination can be mitigated by a factor between $\sim 1.5$ ($|\theta_\textrm{b}|\le$~7\fdg 5) and $\sim 10$ ($|\theta_\textrm{b}|\le$~22\fdg 5). The latter limit is achieved with a mere 5\% loss of data for $|\vec{b}|_\textrm{max}\simeq30$~m. And discarding antenna spacings with lengths $|\vec{b}|\gtrsim 30$~m does not mitigate foreground contamination any further and would only lead to loss of sensitivity as the fraction of discarded baselines increases from $\sim 5$\% to $\sim 25$\%. On the other hand, a bright point source at the same location will give rise to foreground contamination even on longer antenna spacings. Such cases will necessitate discarding more or, in the worst case, all of the eastward oriented antenna spacings.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_nside_64_Tsys_95.0K.eps}
\caption{Drop in foreground contamination in the {\it EoR window}, and loss of data for the {\it off--zenith} pointing as a function of discarded baselines. Eastward baselines with varying degrees of directedness -- $|\theta_\textrm{b}|<$~7\fdg 5 (solid circles), $|\theta_\textrm{b}|<$~15\arcdeg~ (solid squares), and $|\theta_\textrm{b}|<$~22\fdg 5 (solid stars) -- and lengths $|\vec{b}| \le |\vec{b}|_\textrm{max}$ ($x$--axis) are discarded. Loss of data (dotted lines) is measured by discarded baselines as a fraction of the total number for the corresponding cases. Foreground contamination in the {\it EoR window} (solid lines) drops by a factor $\sim$1.5 ($|\theta_\textrm{b}|\le$~7\fdg 5) to $\sim$10 ($|\theta_\textrm{b}|\le$~22\fdg 5). The latter limit can be achieved with a mere 5\% loss of data at $|\vec{b}|_\textrm{max}\simeq30$~m, and discarding longer baselines ($|\vec{b}|\gtrsim 30$~m) has no effect in further reducing foreground contamination. \label{fig:screening}}
\end{figure}

@@

In conclusion, we find that inclusion of emission models, both diffuse and compact, all the way to the horizon is essential to explaining the observed power spectrum. We also provide a simple and effective tool based on the delay spectrum technique to mitigate foreground contamination by an order of magnitude in EoR data analysis by discarding or down--weighting antenna pairs most affected by foreground contamination as a function of antenna spacing vector and time of observation, with negligible loss of sensitivity.
